---
title: "Project"
author: "Diana Fay Arbas"
date: "2023-06-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Team 1 Final Project

First we will import the dataset and describe characteristics such as dimensions, data types, file types, and import methods used.

```{r}

# Import dataset
sleep <- read.csv("sleep.csv")
head(sleep)

```

```{r}

# Get the dimensions of the dataset
dim(sleep)

```

There are 374 rows and 13 columns in the sleep dataset.


```{r}

# Display the structure of the sleep dataset
# including variable names, their data types, and some example values
str(sleep)

```
The target variable is sleep disorder, which is categorical. Other categorical variables are gender, and occupation. The ordinal variables are quality of sleep, stress level, and BMI category. The numerical variables are age, sleep duration, physical activity level, blood pressure, heart rate, and daily steps. 

Now we will check for missing data.

```{r}

# Check for null values
null_values <- is.na(sleep)
sum(null_values)

```

There are no missing values in this dataset.


