---
title: "Project"
author: "Diana Fay Arbas"
date: "2023-06-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Team 1 Final Project

First we will import the dataset and describe characteristics such as dimensions, data types, file types, and import methods used.

```{r}

# Import dataset
sleep <- read.csv("sleep.csv")
head(sleep)

```

```{r}

# Get the dimensions of the dataset
dim(sleep)

```

There are 374 rows and 13 columns in the sleep dataset.


```{r}

# Display the structure of the sleep dataset
# including variable names, their data types, and some example values
str(sleep)

```
The target variable is sleep disorder, which is categorical. Other categorical variables are gender, and occupation. The ordinal variables are quality of sleep, stress level, and BMI category. The numerical variables are age, sleep duration, physical activity level, blood pressure, heart rate, and daily steps. 

Now we will check for missing data.

```{r}

# Check for null values
null_values <- is.na(sleep)
sum(null_values)

```

There are no missing values in this dataset.

Let's provide measures of centrality and distribution with visualizations on the following predictors: sleep duration, quality of sleep, and heart rate. 

```{r}

# Sleep duration descriptive statistics
summary(sleep$Sleep.Duration)

```

The mean sleep duration is 7.132, and the median sleep duration is 7.2.

```{r}

hist(sleep$Sleep.Duration, main = "Histogram of Sleep Duration", xlab = "Sleep Duration")

```
```{r}

# Quality of sleep descriptive statistics
summary(sleep$Quality.of.Sleep)

```

On a scale of 1-10, the mean quality of sleep was 7.3 and the median was 7.

```{r}

hist(sleep$Quality.of.Sleep, main = "Histogram of Quality of Sleep", xlab = "Quality of Sleep")

```
```{r}

# Heart rate descriptive statistics
summary(sleep$Heart.Rate)

```

The mean heart rate is 70.17, and the median heart rate is 70.

```{r}

hist(sleep$Heart.Rate, main = "Histogram of Heart Rate", xlab = "Heart Rate")

```


